how to \texttt{git bisect} when your test suite is having a bad day

### 2020-03-30 bisect/paper/paper.tex
finish writing body of paper

\usepackage{yfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
%\usepackage{mathpartir}
\usepackage[colorlinks=true,
citecolor=citec,
linkcolor=linkc,
urlcolor=urlc,
]{hyperref}
\usepackage{url}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{harmony}
\usepackage{marvosym}
\usepackage{multirow}
\usepackage{xspace}
\usepackage[nameinlink,nosort]{cleveref}
\usepackage{xeCJK}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utopia]{mathdesign}
\usepackage{natbib}
\usepackage{ulem}
\usepackage[mathcal]{euscript}
\usepackage[linesnumbered,ruled]{algorithm2e}

@@

@@

\keywords pdflatex
% bracket, groove, in, jumps, the

@@

Ok, but what if the bug reprodues nondeterministically with some probability $p<1$.
You can't even pinpoint the bug at some $c_{b}$ for sure anymore;
you can at best know
that it \textasciitilde{}prooobably\textasciitilde~won't reproduce in any $c_{i<b}$,
with some confidence $\confidents$.
Now evaluate your strategy by the expected number of steps to achieve, let's say for tradition's sake, $\confidents \ge 0.99999$.
Is it still optimal to bisect at the midpoint?%
\footnote{Spoiler: No, or I wouldn't have bothered with this paper.}
What would be a better strategy, as a function of $p$?

@@

Ok, so let's model the problem as a sequence of steps on a probability distribution function
(henceforth, PDF; and also, CDF for the cumulative kind).
Initially, $\pdf(i) = 1/n$ for all $0 \le i < n$.
When the bug reproduces at some commit $b$,
you're certain no $c_{i>b}$ introduced the bug, so each $\pdf(i>b)$ becomes $0$,
and each $\pdf(i\le b)$ renormalizes by a factor of $n/b$, to preserve the $\int \pdf(i) \text{d} i = 1$ invariant.%
\footnote{Implemented as {\tt fn adjust\_pdf\_bug\_repros()} in the simulator.}

In the deterministic case ($p=1$),
the symmetric thing happens when the test passes at some $c_j$: $\pdf(i\le j) = 0$.
But when $p<1$, we must generalized it \cite{mario3}. Here's Bayes's rule:
\[
P(A|B) = \frac{P(B|A) \times P(A)}{P(B)}
\]

@@

Let's do a nice concrete example.
Say $n=16$, the test passes at $j=7$, and then the bug repros at $j=11$.
In the deterministic case, all the probability mass will be concentrated uniformly in the range $[8,11]$.
However, if the bug repros only half the time,
%the renormalization factors around $c_7$ come out to $2/3$ and $1/3$
$\renorm{i \le 7} = 2/3$
and
$\renorm{i > 7} = 4/3$,
and we get probability mass scattered all the way down to $c_0$,
as shown in \Cref{fig:example}(a).
Yuck, someone clean that up!

Now let's say the test passes at $j=9$, then at $j=10$.
\Cref{fig:example}(b) shows the updated PDFs/CDFs:
for $p=1$, this pinpoints the bug at $j=11$, and the search is over.
But for $p=0.5$, there's still 2/3 odds we'd be wrong!
In fact, from here it takes 18 further probes at $j=10$ until we are at least five 9's confident that $c_{11}$ is the culprit.%
\footnote{See {\tt fn test\_figure\_1()}!}
Bayes's rule gonna get ya.

@@

\begin{figure}[t]
	\begin{tabular}{c}
	\includegraphics[width=0.45\textwidth]{example.pdf}
	\\
		(a) After two tests.
		\\
		\\

	\includegraphics[width=0.45\textwidth]{example2.pdf}
	\\
		(b) After ``classical`` search terminates.
	\end{tabular}
	%\caption{Example deterministic and nondeterministic PDFs and CDFs.}
	\caption{Example \{,non\}deterministic \{P,C\}DFs.}
	\label{fig:example}
\end{figure}

@@

\section{prior work}

@@

Dropbox has a CI service called Athena \cite{athena}
which automatically searches for flaky tests, when otherwise no bug exists.
In this case the goal is to keep the build green, but if you consider the flaky test itself to be the bug, it's the same problem.%
\footnote{Incidentally, the symmetric case
-- where a bug repros with $p=1$, but the test also flakes with some $q<1$ --
is also the same problem.}
Athena's ``deflakes'' the test at each commit by running 10 times,
treating the combined result as ``basically as good as $p=1$'',
and then runs an otherwise classical binary search.%
\footnote{Upcoming, I will call this strategy \textsf{Human-Mistrustful}$(10)$.}
In this setting, $p$ is not known in advance, so using Bayes's rule is off the table.
But I will show that even without access to the PDF, a better strategy exists. % at least by my definition

@@

\section{strategies}

Ok, so how do we
%cope with the fact that we'll never have a proper lower bound?
%How do we
make progress,
i.e., concentrate probability mass til there's $\confidents$ of it in one place,
as quickly as possible?
%In principle, testing anywhere in the range $[0,b)$,
%where $b$ is the earliest known buggy commit,
%will make progress,
%
%% ghhh this sentence only feels good if i know what i'm talking about
Let's deconstruct the motives of classical binary search.
Let $c_b$ be the earliest known buggy commit, and $c_a$ be the latest known safe commit.
In Determinism World, bisecting at $c_{(a+b)/2}$
minimizes the worst case maximum range remaining,
as the two possible outcome ranges are the same length.
%But in Nondeterminism World,
%the worst case is a big constant times linear:
%an adversarial RNG god could pass your test a thousand times on $c_{n-1}$, then fail,
%then pass it a thousand times on $c_{n-2}$, then fail, and so on,
%before ultimately arriving at buggy commit $c_0$.
%So thinking about the worst case is not productive for a sublinear strategy,
%and also,
%we shouldn't be thinking about the range length,
%because a pass is still progress even though $a$ doesn't ever budge,
But in Nondeterminism World, $a$ doesn't ever budge from $0$,
so bisecting at $c_{(0+b)/2}$ will not even terminate.
%will eventually fail to make progress.
Sure, hitting the PDF with
%Bayes's rule at $b/2$
$\renorm{b/2}$
will always move {\it some} mass rightward,
but once five 9's of it is already over there,
%to the right of that point,
it can't concentrate it onto one point.
So let's not think about the range.

\subsection{bisect probability mass}

Another way to think about Determinism World is that the $c_{(a+b)/2}$ bisect point
% maximizes the expected probability mass that moves from one side to the other.
cuts the probability mass, rather than the range, in half,
% How about in Nondeterminism World,
% we ignore the range length, and instead bisect at the midpoint of probability mass,
i.e., $\mathsf{max}_j (i)$ s.t. $\cdf(j) \le 0.5)$.
This approach
fits the ``binary search'' spirit,
and will terminate in Nondeterminism World,
i.e.,
if $b$ is the solution, it converges to repeatedly probing $b-1$,
so it can pass any fixed $\confidents$ threshold.
%(which is more than you can say for the range-based $c_{(0+b)/2}$).
%It is intuitively obvious that this will make progress faster than bisecting at $c_{(b+0)/2}$,
%But is it fast?
% However, it doesn't maximize either the minimum or the expected amount of mass that moves:
% when $p<1$, it's intuitively apparent that this CDF threshold should actually be less than $0.5$.%
%\footnote{I did some pen-and-paper algebra to find a closed form for this,
%but I actually trust the simulator more than myself at this point,
%so I stopped spending brains on this part of the problem. I kind of regret that.}
%
%Is maximizing the expected moved mass as a function of $p$ the fastest we can go,
%and if so, why?
But is $0.5$ still the best bisect point even when $p<1$?
I wondered if this could be expressed as the amount of probability mass moved from one side to the other,
i.e., $\sum_i \mathsf{abs}(\pdf(i) - \pdf'(i))$.
In the case where the bug repros this is:
\[
	p\cdf(j) \times (1-\cdf(j))
\]
and in the case where the test passes:
\[
	(1-p\cdf(j)) \times \cdf(j) \times (1-\renorm{i\le j})
\]
which, surprisingly, simplifies to exactly the same thing as the bug repros case.
The maximum occurs where $\partial/\partial \cdf(j)$ is 0,
which turns out to be at $\cdf(j)=0.5$ after all, and independent of $p$.%
\footnote{I also worked out the {\it minimum} moved mass between the pass and fail case,
which is almost always the pass case,
and comes out to
%$\frac{p\cdf(j)(1-\cdf(j))}{1-p\cdf()}$
$\frac{\cdf(j)(1-\cdf(j))}{1/p-\cdf(j)}$,
which experiences its maximum at
$\frac{1-\sqrt{1-p}}{p}$ (thanks wolframalpha).
But this is worst-case thinking, which doesn't seem appropriate.
Let's have some optimism!}
But it's not clear that the amount of mass moved
necessarily corresponds to reaching $\confidents$ the fastest.
%It just sounds kinda nice?
%What is the best threshold, perhaps as a function of $p$, and why?

@@

\subsection{bisect entropy}

@@

A PDF's information content is measured in entropy:
$\entropy = \sum_i \pdf(i)\mathsf{ln}(\pdf(i))$.
Alone, this value is fairly meaningless, but to compare them,
a spikier PDF has lower entropy than a flatter one.
Indeed,
when the search terminates in Determinism World, $\entropy = 0$.
I thought for a while about how to link ``minimum expected entropy''
to the stated goal %in Nondeterminism World:
of
%minimizing the expected steps until
$\confidents = \mathsf{max}_i (\pdf(i)) \ge 0.99999$,
but couldn't really formalize the idea.
The goal of five 9's is fairly arbitrary anyway,
and also not necessarily stable under the entropy measurement,
since it doesn't care how the remaining $0.00001$ is distributed.%
\footnote{See {\tt fn test\_entropy\_stability()}.}

This strategy is expensive to compute.
Whereas bisecting at some fixed threshold of probability mass merely requires a $O(\mathsf{log}  n)$ query of the CDF,
computing the expected entropy is $O(n)$ for {\it each} bisect point.
A Very Smart Math Friend of mine \cite{primer} analyzed the easy case of the initial test,
when the PDF is still uniform (i.e., $\forall i, \cdf(i) = i$),
and found the closed form
\[
	\entropy_{j,0} = p \mathsf{ln}(j)
	+
	(1-pj) \mathsf{ln} (1-pj)
	-
	j(1-p) \mathsf{ln} (j-p)
\]
whose derivative,
in his words,
``looks like a giant mess that does not admit an analytic solution for $j$.''

@@

\subsection{bisect randomly}

@@

You can tell at this point I'm starting to get statistics fatigue.

\subsection{human}

Humans are known to be less patient than computers (citation: see previous sentence).
If a human is unwilling to compute $\renorm{} \circ \pdf$ by hand every step,
or, more prohibitively,
simply doesn't know $p$ in advance so can't compute $\renorm{}$ at all,
what should they do?

Think back to Athena from the prior work section.
The human

% parameterize over 10
% what should the human do when they think they're done? -> probe repeatedly at b-1
% what should they do when

Regardless, it's easy to

@@

\section{simulating it}

@@

To save you some page-turning, here's the source code link again:
\url{https://github.com/bblum/sigbovik/blob/master/bisect}.
A {\tt BisectStrategy}
implements a function over $p$, the PDF, and/or its own internal state
to choose the next bisect point.
The {\tt SimulationState} is initialized with fixed $n$ and $p$,
and repeatedly invokes a given {\tt BisectStrategy},
hitting the PDF with Bayes's rule accordingly,
until $\confidents \ge 0.99999$.
I provide implementations for all the strategies from the previous section.
It's written in Rust so it's self-documenting.

I learned a lot about floating point.
I don't mean like NaNs and mantissa bits,
I mean that when your problem is fundamentally a long multiplication sequence of increasingly high rationals,
your precious $\cdf(n)$ inevitably drifts ever farther from 1, and you start drowning in imprecision \cite{drowningpoint}.
I suppose I could have used the {\tt BigRational} library,
but I chose Rust over Haskell for this so I could {\it avoid} runaway memory consumption,
so I resigned myself to explicitly renormalizing my PDFs by $1/\cdf(n)$ each time I updated them.
After that, I was able to write some assertions
to bound the imprecision drift within $k(\text{\tt std::f64::EPSILON})$ \cite{epsilon}.
But yeah, I definitely spent some sanity points writing a function named {\tt fn assert\_kinda\_equals\_one(\&self)}.

@@

\section{evaluation}

@@

I'm not ashamed to admit I left a lot of loose ends in this paper.
The biggest open questions for me are:
\begin{enumerate}
	\item What relates $p$ to {\sf cdfbisect}'s optimal argument?
	\item Why is {\sf entropy} {\it so tantalizingly close} to {\sf cdfbisect}? Is there some way to show they're the same?
\end{enumerate}
I have also prepared a set of backup questions in case those aren't enough for you:
\begin{enumerate}
	\setcounter{enumi}{3}
	\item If you don't know $p$ in advance, is there an efficient way to multiplex the search with figuring it out on the fly?
	\item What if false positives are also possible, with some $q \ne p$?
	\item How do you account for the cost of {\tt git checkout} being nonzero (i.e., having to recompile your tests on new code)?
		Clearly {\sf human-mistrustful} becomes better. Where is the break-even point?
\end{enumerate}

>>>>

\usepackage{yfonts}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
%\usepackage{mathpartir}
\usepackage[colorlinks=true,
citecolor=citec,
linkcolor=linkc,
urlcolor=urlc,
]{hyperref}
\usepackage{url}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{wasysym}
\usepackage{harmony}
\usepackage{marvosym}
\usepackage{multirow}
\usepackage{xspace}
\usepackage[nameinlink,nosort]{cleveref}
\usepackage{mathspec}
\usepackage{xeCJK}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[utopia]{mathdesign}
\usepackage{natbib}
\usepackage{ulem}
\usepackage[mathcal]{euscript}
\usepackage[linesnumbered,ruled]{algorithm2e}

@@

\renewcommand{\figurename}{figure}

\crefformat{section}{#2{section~#1}#3}
\Crefformat{section}{#2{Section~#1}#3}
\Crefformat{figure}{#2{Figure~#1}#3}
\crefformat{figure}{#2{figure~#1}#3}
%\Crefformat{Figure}{#2{figure~#1}#3}

@@

\keywords pdflatex, cdflatex
% bracket, groove, in, jumps, the

@@

Ok, but what if the bug reprodues nondeterministically with some probability $p<1$.
You can't even pinpoint the bug at some $c_{b}$ for sure anymore;
you can at best know
that it \textasciitilde{}prooobably\textasciitilde~won't reproduce in any $c_{i<b}$,
with some confidence $\confidents$.
Now evaluate your strategy by the expected number of steps to achieve, let's say for tradition's sake, $\confidents > 0.99999$.
Is it still optimal to bisect at the midpoint?%
\footnote{Spoiler: No, or I wouldn't have bothered with this paper.}
What would be a better strategy, as a function of $p$?

@@

Ok, so let's model the problem as a sequence of steps on a probability distribution function
(henceforth, PDF; and also, CDF for the cumulative kind).
Initially, $\pdf(i) = 1/n$ for all $0 \le i < n$.
When the bug reproduces at some commit $b$,
you're certain no $c_{i>b}$ introduced the bug, so each $\pdf(i>b) \leftarrow 0$
and each $\pdf(i\le b)$ renormalizes by a factor of $n/b$, to preserve the
%$\int \pdf(i) \text{d} i = 1$
$\sum_i \pdf(i) = 1$
invariant.%
\footnote{Implemented as {\tt fn adjust\_pdf\_bug\_repros()} in the simulator.}

In the deterministic case, $p=1$,
the symmetric thing happens when the test passes at some $c_j$: $\pdf(i\le j) \leftarrow 0$.
But when $p<1$, we must generalized it \cite{mario3}. Here's Bayes's rule:
\[
P(A|B) = \frac{P(B|A) \times P(A)}{P(B)}
\]

@@

Let's do a nice concrete example.
Say $n=16$, the test passes at $j=7$, and then fails at $j=11$.
In the deterministic case, all the probability mass will be concentrated uniformly in the range $[8,11]$.
However, if the bug repros only half the time,
%the renormalization factors around $c_7$ come out to $2/3$ and $1/3$
$\renorm{i \le 7} = 2/3$
and
$\renorm{i > 7} = 4/3$,
and we get probability mass scattered all the way down to $c_0$,
as shown in \cref{fig:example}(a).
Yuck, someone clean that up!

Now let's say the test passes at $j=9$, then at $j=10$.
\cref{fig:example}(b) shows the updated PDFs/CDFs:
for $p=1$, this pinpoints the bug at $j=11$, and the search is over.
But for $p=0.5$, there's still 2/3 odds we'd be wrong!
In fact, from here it takes 18 further probes at $j=10$ until we are at least five 9's confident that $c_{11}$ is the culprit.%
\footnote{See {\tt fn test\_figure\_1()}!}
Bayes's rule gonna get ya.

@@

\begin{figure}[t]
	\begin{tabular}{c}
	\includegraphics[width=0.45\textwidth]{example.pdf}
	\\
		(a) after two tests
		\\
		\\

	\includegraphics[width=0.45\textwidth]{example2.pdf}
	\\
		(b) after ``classical`` search terminates
	\end{tabular}
	%\caption{Example deterministic and nondeterministic PDFs and CDFs.}
	\caption{example \{,non\}deterministic \{P,C\}DFs}
	\label{fig:example}
\end{figure}

@@

\section{prior work}
\label{sec:prior}

@@

Dropbox has a CI service called Athena \cite{athena}
which automatically searches for flaky tests. %, when otherwise no bug exists.
In this case the goal is to keep the build green,
but if you consider the flaky test itself to be the bug, it's the same problem.%
\footnote{Incidentally, the symmetric case
-- where a bug repros with $p=1$, but the test also flakes with some $q<1$ --
is also the same problem.}
Athena's ``deflakes'' the test at each commit by running 10 times,
treating the combined result as ``basically as good as $p=1$'',
and then runs an otherwise classical binary search.
%\footnote{Upcoming, I will call this strategy \textsf{Human-Mistrustful}$(10)$.}
In this setting, $p$ is not known in advance, so using Bayes's rule is off the table.
But I will show that even without access to the PDF, a better strategy exists. % at least by my definition

@@

\section{strategies}
\label{sec:strats}

Ok, so how do we
%cope with the fact that we'll never have a proper lower bound?
%How do we
make progress,
i.e., concentrate probability mass til there's $\confidents$ of it in one place,
as quickly as possible?
%In principle, testing anywhere in the range $[0,b)$,
%where $b$ is the earliest known buggy commit,
%will make progress,
%
%% ghhh this sentence only feels good if i know what i'm talking about
Let's deconstruct the motives of classical binary search.
Let $c_b$ denote the earliest known buggy commit, and $c_a$ be the latest known safe commit.
In Determinism World, bisecting at $c_{(a+b)/2}$
minimizes the worst case maximum range remaining,
as the two possible outcome ranges are the same length.
%But in Nondeterminism World,
%the worst case is a big constant times linear:
%an adversarial RNG god could pass your test a thousand times on $c_{n-1}$, then fail,
%then pass it a thousand times on $c_{n-2}$, then fail, and so on,
%before ultimately arriving at buggy commit $c_0$.
%So thinking about the worst case is not productive for a sublinear strategy,
%and also,
%we shouldn't be thinking about the range length,
%because a pass is still progress even though $a$ doesn't ever budge,
But in Nondeterminism World, $a$ doesn't ever budge from 0,
so bisecting at $c_{(0+b)/2}$ will not even terminate.
%will eventually fail to make progress.
Sure, hitting the PDF with
%Bayes's rule at $b/2$
$\renorm{b/2}$
will always move {\it some} mass rightward,
but once five 9's of it is already over there,
%to the right of that point,
it can't concentrate it onto one point.
So let's not think about the range.

\subsection{bisect probability mass}
\label{sec:cdfbisect}

Another way to frame it is that the $c_{(a+b)/2}$ bisect point
% maximizes the expected probability mass that moves from one side to the other.
cuts the probability mass, rather than the range, in half,
% How about in Nondeterminism World,
% we ignore the range length, and instead bisect at the midpoint of probability mass,
i.e., $\mathsf{max}_j (i)$ s.t. $\cdf(j) \le 0.5)$.
This approach
fits the ``binary search'' spirit,
and will terminate in Nondeterminism World,
i.e.,
if $b$ is the solution, it converges to repeatedly probing $b-1$,
so it can pass any fixed $\confidents$ threshold.
%(which is more than you can say for the range-based $c_{(0+b)/2}$).
%It is intuitively obvious that this will make progress faster than bisecting at $c_{(b+0)/2}$,
%But is it fast?
% However, it doesn't maximize either the minimum or the expected amount of mass that moves:
% when $p<1$, it's intuitively apparent that this CDF threshold should actually be less than $0.5$.%
%\footnote{I did some pen-and-paper algebra to find a closed form for this,
%but I actually trust the simulator more than myself at this point,
%so I stopped spending brains on this part of the problem. I kind of regret that.}
%
%Is maximizing the expected moved mass as a function of $p$ the fastest we can go,
%and if so, why?
But is $0.5$ still the best bisect point even when $p<1$?
I wondered if this could be expressed as the amount of probability mass moved from one side to the other,
i.e., $\sum_i \mathsf{abs}(\pdf(i) - \pdf'(i))$.
In the case where the bug repros this is:
\[
	p\cdf(j) \times (1-\cdf(j))
\]
and in the case where the test passes:
\[
	(1-p\cdf(j)) \times \cdf(j) \times (1-\renorm{i\le j})
\]
which, surprisingly, simplifies to exactly the same thing as the bug repros case.
The maximum occurs where $\partial/\partial \cdf(j)$ is 0,
which turns out to be at $\cdf(j)=0.5$ after all, and independent of $p$.%
\footnote{I also worked out the {\it minimum} moved mass between the pass and fail case,
which is almost always the pass case.
It comes out to
%$\frac{p\cdf(j)(1-\cdf(j))}{1-p\cdf()}$
$\frac{\cdf(j)(1-\cdf(j))}{1/p-\cdf(j)}$,
which experiences its maximum at
$\frac{1-\sqrt{1-p}}{p}$ (thanks wolframalpha).
But this is worst-case thinking, which doesn't seem appropriate.
Let's have some optimism!}
But it's not clear that the amount of mass moved
necessarily corresponds to reaching $\confidents$ the fastest.
%It just sounds kinda nice?
%What is the best threshold, perhaps as a function of $p$, and why?
I show my work in \cref{fig:let-me-shown-that-my-work},
because that's what they taught me to do in high school.

\begin{figure}[t]
	\begin{tabular}{c}
		\includegraphics[width=0.45\textwidth]{horses.jpg}
	\end{tabular}
	\caption{derivation(?) of optimal(??) bisect point}
	\label{fig:let-me-shown-that-my-work}
\end{figure}

@@

\subsection{bisect entropy}
\label{sec:entropy}

@@

A PDF's information content is measured in entropy:
$\entropy = \sum_i \pdf(i)\mathsf{ln}(\pdf(i))$.
Alone, this value is fairly meaningless, but to compare them,
a spikier PDF has lower entropy than a flatter one.
Indeed,
when the search terminates in Determinism World, $\entropy = 0$.
I thought for a while about how to link ``minimum expected entropy''
to the stated goal %in Nondeterminism World:
of
%minimizing the expected steps until
$\confidents = \mathsf{max}_i (\pdf(i)) > 0.99999$,
but couldn't really formalize the idea.
The goal of five 9's is fairly arbitrary anyway,
and also not necessarily stable under the entropy measurement,
since it doesn't care how the remaining $0.00001$ is distributed.%
\footnote{See {\tt fn test\_entropy\_stability()}.}

This strategy is expensive to compute.
Whereas bisecting at some fixed threshold of probability mass merely requires a $O(\mathsf{log}  n)$ query of the CDF,
computing the expected entropy is $O(n)$ for {\it each} bisect point.
A Very Smart Math Friend of mine \cite{primer} analyzed the easy case of the initial test,
when the PDF is still uniform (i.e., $\forall i, \cdf(i) = i$),
and found the closed form:
\[
	\entropy_{0,j} = p \mathsf{ln}(j)
	+
	(1-pj) (\mathsf{ln} (1-pj))
	-
	j(1-p) (\mathsf{ln} (j-p))
\]
whose derivative,
in his words,
``looks like a giant mess that does not admit an analytic solution for $j$.''

@@

\subsection{bisect randomly}
\label{sec:random}

@@

You can tell at this point I'm starting to get statistics fatigue.
Not unprecedented in these hallowed proceedings \cite{batch-myass}.

\subsection{human}
\label{sec:human}

Humans are known to be less patient than computers (citation: \cref{sec:random}).
If a human is unwilling to compute $\renorm{} \circ \pdf$ by hand every step,
or, more prohibitively,
simply doesn't know $p$ in advance so can't compute $\renorm{}$ at all,
what should they do?
Let's say a ``human strategy'' is any that doesn't use the PDF/CDF when deciding where to bisect.%
%i.e., one that's easy for a human to do without bothering with Bayes's rule.%
\footnote{In fact {\sf rand-uniform} from last section counts as a human strategy. Did you bring your dice?}

The most straightforward thing for a human to do is to assume $p=1$ until they encounter evidence to the contrary.
They'll conduct a normal binary search, and when they think they've found the bug at $c_b$,
they'll just keep probing $c_{b-1}$ until enough probability mass (invisible to them)
moves to $b$ and the simulator says they can stop.
If this instead repros the bug at $c_{b-1}$ after all, the human gets confused!
They'll assume the bug might have been present as early as $c_0$,
forgets their lower bound, and begins binary searching anew in the range $[0,b-1]$.
It's easy to see this makes progress.
Let's call this the ``confused, forgetful human'' strategy.

Another confused human variant is for them to remember their previous lower bounds.
After all, just because their best lower bound $b-1$ was contradicted
doesn't say anything about some earlier bound $a<b-1$ they might have observed.
So this human will maintain a stack of lower bounds, and when $b-1$ is contradicted,
they'll pop off the stack and resume searching in $[a,b-1]$ instead.
Let's call this the ``confused human who remembers''.
%
Incidentally, while implementing this strategy, I accidentally wrote the comment,
\begin{center}
	\begin{tabular}{l}
		{\tt // the human, who now has perfect memory,} \\
		{\tt // walks backwards in time}
	\end{tabular}
\end{center}
and giggled a lot to myself as I imagined {\birbaslo sigbovik}~fanfiction.

Finally, let's imagine the human knows maybe something is up with this $p$ stuff in advance. %before they start searching.
Wishing to carry on classically as though $p$ were 1,
they'll try to emulate $p$ being higher by retrying any passing test ``just to be sure'', before moving on.
If they retry the test $r$ times,
the probability it fails given the bug is present is then $1-(1-p)^{1+r}$.
%In general, for $r$ retries, that's $1-(1-p)^{r+1}$.
But it comes at the cost of $r$ more steps for each new lower bound!
As before, when this human thinks they're done, they'll repeatedly probe $c_{b-1}$ until contradicted
(and we'll make them forgetful, to keep things simple).
Let's call this the ``suspicious human of $r$ retries''.

@@

\section{simulated it}

@@

To save you some page-turning, here's the source code link again:
\url{https://github.com/bblum/sigbovik/blob/master/bisect}.
A {\tt BisectStrategy}
implements a function over $p$, the PDF, and/or its own internal state
to choose the next bisect point.
The {\tt SimulationState} is initialized with fixed $n$ and $p$,
and repeatedly invokes a given {\tt BisectStrategy},
hitting the PDF with Bayes's rule accordingly,
until $\confidents > 0.99999$.
I provide implementations for all the strategies from the previous section in {\tt src/strategies/}.
It's written in Rust so it's self-documenting.

I learned a lot about floating point.
I don't mean like NaNs and mantissa bits,
I mean that when your project is fundamentally a long multiplication chain of increasingly high rationals,
your precious $\cdf(n)$ inevitably drifts farther from 1 than you can bound with any arbitrary $\epsilon$, and you start drowning in imprecision \cite{drowningpoint}.
I suppose I could have used the {\tt BigRational} library,
but I chose Rust over Haskell for this so I could {\it avoid} runaway memory consumption...
so I resigned myself to explicitly renormalizing my PDFs by $1/\cdf(n)$ each time I updated them.
After that, I was able to write some assertions
to bound the imprecision drift within $k(\text{\tt std::f64::EPSILON})$ \cite{epsilon}.
But yeah, I definitely spent some sanity points writing a function named {\tt fn assert\_kinda\_equals\_one(\&self)}.

@@

\section{experiments}

Let's get right to the good stuff.
\cref{fig:resulce} shows the overall results,
plotting the average performance of each of the strategies from \cref{sec:strats}.

\begin{figure}[ht]
	\begin{tabular}{c}
		\includegraphics[width=0.45\textwidth]{resulce.pdf}
		\\
		(a) full view
		\\
		\\
		\includegraphics[width=0.45\textwidth]{resulce-detail.pdf}
		\\
		(b) detail view of high $p$ range; linear scale
	\end{tabular}
	\caption{perf results.
	how many bisect steps each strategy takes to reach five 9s of confidence} %,
	%simulated on a 1024-long {\tt git} {\tt log}; 65536 trials}
	\label{fig:resulce}
\end{figure}

Each data point is simulated with $n=1024$ and averaged over 65536 trials.
I chose a power of two for $n$ to minimize floating point drift (handwave),
and a multiple of $n$ for trials so that {\tt buggy\_commit = trial \% n} would be uniform.

Of course, each trial takes exponentially
%\footnote{Yes, I am using this word in the literal sense.}
longer and longer as $p \rightarrow 0$, so I didn't bother testing past 0.1.
I also didn't bother measuring execution time,
as prior work has shown that one's personal laptop is a fraught environment for stable performance evaluation \cite{htm},
but it was still very obvious that {\sf entropy} was far, far slower per step than all other strategies.
%I guess I'd say it took minutes to run {\sf cdfbisect} up to $p=0.01$, and hours/overnight for the same for {\sf entropy}.

\subsection{computer strategies}

Minimizing expected entropy
%({\sf entropy}; \Cref{sec:entropy})
turned out to be the best strategy, globally across all $p$.
Bisecting probability mass
%({\sf cdfbisect}; \Cref{sec:cdf})
turns out to be preeetty close,
although its performance varies depending on its bisect-point parameter $j$.
Obviously, when $p=1$, $j$ should be 0.5 to reproduce classical binary search,
but at around $p=0.8$, $j=0.333$ overtakes it.
Ultimately at $p=0.1$, the former is about 4\% slower compared to the latter and to entropy,
contradicting the ``independent of $p$'' hypothesis from \cref{sec:cdfbisect}.
I investigated a little more to see how $p$ affected the optimal $j$, testing various $j$s in increments of 0.01:
\begin{center}
	\begin{tabular}{c|cc}
		$p$ & best $j$ & {\sf cdfbisect}$(j)$ \\
		\hline
		1.0 & 0.50 & 10     \\
		0.9 & 0.45 & 17.4   \\
		0.8 & 0.41 & 21.9   \\
		0.7 & 0.39 & 27.5   \\
		0.6 & 0.40 & 34.6   \\
		0.5 & 0.36 & 44.1   \\
		0.4 & 0.35 & 58.3   \\
		0.3 & 0.34 & 81.6   \\
		0.2 & 0.36 & 127.9  \\
		0.1 & 0.35 & 266.6  \\
	\end{tabular}
	%\begin{tabular}{l|c ccccc ccccc}
	%	$p$				& 1 & 0.9 & 0.8 & 0.7 & 0.6
	%						& 0.5 & 0.4 & 0.3 & 0.2 & 0.1 \\
	%	\hline
	%	best $j$			& 0.5 & 0.45 & 0.41 & 0.39 & 0.40
	%						& 0.36 & 0.35 & 0.34 & 0.36 & 0.35 \\
	%	best {\sf cdfbisect}$(j)$	& 10 & 17.4 & 21.9 & 27.5 & 34.6
	%						& 44.1 & 58.3 & 81.6 & 127.9 & 266.6
	%\end{tabular}
\end{center}
65536 trials here wasn't really enough to get variance under control,
but it definitely seems to converge towards 0.333 or maybe 0.35 or something as $p \rightarrow 0$.
Open question why for warp zone, I guess.

{\sf rand-uniform} is obviously terrible but I threw it in there for kicks.
{\sf rand-cdf} is not so terrible, doing even better than all the human strategies.

\subsection{human strategies}

I investigated how the suspicious human's number-of-retries parameter $r$ affects their performance.
In this notation, $r=1$ means they retry once whenever the test passes, meaning two test runs on every passing commit.
$r=0$ is equivalent to the forgetful strategy, and $r=9$ is what Athena does (\cref{sec:prior}).%
% actually it's an enhanced version of athena so that it will terminate

\cref{fig:resulce} was getting a little crowded so I've plotted all the human-friendly strategies on \cref{fig:human}.
It's pretty self-explanatory imo.

\begin{figure}[t]
	\begin{tabular}{c}
		\includegraphics[width=0.45\textwidth]{resulce-human.pdf}
	\end{tabular}
	\caption{version of \cref{fig:resulce} for human strategies}
	\label{fig:human}
\end{figure}

One thing that surprised me was that the human who forgets all their past known lower bounds when {\it any} of them is contradicted
performed better than the human who tried
to rewind to their last not-yet-contradicted lower bound.
I guess there's a moral here about not clinging to the past, moving on with life, or something.

\subsection{why five 9s?}

Let's talk about the arbitrary confidence threshold.
There's a lot of knobs to tweak here and only so much time til April 1 (not to mention reader patience),
but I did at least want to show what happens if you terminate the search when you're only 50\% confident.
Compared to \cref{fig:resulce},
the strategies all terminate in about half as many steps, with the same relative ranking and break-even points.
What's more interesting is the number of times they were wrong.

I had the simulator refuse to count a wrong result, %towards the 65536 trial count,
restarting the search until it correctly identified the buggy commit 65536 times,
and counting the number of restarts.
With $\confidents > 0.99999$ this was not very interesting:
the strategies collectively made 0 errors 81\% of the time, and were wrong more than twice only 0.6\%.
But $\confidents > 0.5$ was a different story.
As \cref{fig:wrongs} shows, most strategies
-- even the comically slow {\sf rand-uniform} --
converged to being wrong about half the time, i.e., 65536 wrongs per 65536 rights.
{\sf athena} was most resilient, at least for higher $p$.

\begin{figure}[t]
	\begin{tabular}{c}
		\includegraphics[width=0.45\textwidth]{wrongs.pdf}
	\end{tabular}
	\caption{how often each strategy was wrong, when terminating with only 50\% confidence.
	the dashed line marks the approximate convergence point at 65536, equal to the number of right answers}
	\label{fig:wrongs}
\end{figure}

{\sf confused-human-who-remembers}, alone, did not converge.
They became more and more unable to cope with the repro rate approaching 0,
utlimately being wrong 8 times as often as they were right.
Oh, human... please learn to let go and love yourself for who you are.
% japanese speakers should be able to hear the untranslated "kagayaiteru" here
There's a better world out there, a sparkling radiant future waiting just for you!

@@

I'm not ashamed to admit I left a lot of loose ends in this paper.
The biggest open questions for me are:
\begin{enumerate}
	\item What relates $p$ to {\sf cdfbisect}'s optimal argument?
	\item Why is {\sf entropy} {\it so tantalizingly close} to {\sf cdfbisect}? Is there some way to show they're the same?
\end{enumerate}
I have also prepared a set of backup questions in case those aren't enough for you:
\begin{enumerate}
	\setcounter{enumi}{3} % intentional
	\item If you don't know $p$ in advance, is there an efficient way to multiplex the search with figuring it out on the fly?
	\item What if false positives are also possible, with some flake rate $q \ne p$?
		% TODO make this convey the sigbovik spirit more
	\item What if the cost of {\tt git} {\tt checkout} is nonzero (e.g., having to recompile your tests on new code)?
		Clearly {\sf human-mistrustful} becomes better.
		% Where is the break-even point?
		What about some wacky hybrid {\sf cdfbisect-mistrustful}?
\end{enumerate}
